---
title: "Day 4 Notes and Comp Time"
author: "Colleen McCamy"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)

```

# Brain warm-up - creating a for loop

```{r}

# total time for any two quater miles; find all the times for any consecutive two quarters

quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

for (i in seq_along(quarter_splits)){
  print(quarter_splits[i] + quarter_splits [i+1])
}


# need to specific where the i is indexing from - thus include the vector[i]

```

# Intro to functions
making a function to double the user input

```{r}

double_this <- function(my_value){
  print(2 * my_value)
}

# now using the function lik

double_this(my_value = 14)

double_this(200)

# body of function tell it what to do

# don't need to specify argument name but need to have the same order

double_this(c(1, 2, 10, 100, 175))

#can use a vector too
```


Write a function that asks for a users name and age and returns "I'm [name] and I am [age] years old

```{r}

age_name <- function(name, age) {
  print(paste("I'm", name, "and I am", age, "years old." ))
}

age_name(name = "Colleen", age = 26)

# need to use paste where there are different types of data

```

# Looking at function documentation
Docs contains: 
- Title of function
- what the function name is and package in the top left
- description = brief summary
- usage = an idea of what the actual structure is of the function and what arguments are required for the function
- arguments = idea of default settings that are included
- Value = ideas of what outputs are from model functions; can show what all the different outputs are stored in that list

# Function with a conditional

```{r}

convert_animal_age <- function(animal, human_age){
  if (animal == "dog"){
    print(human_age *7)
  } else if (animal == "goat"){
    print(human_age * 4.7)
  } else {
    print("Please input a dog or goat")
    }
}

convert_animal_age(animal = "dog", human_age = 12)
convert_animal_age(animal = "goat", human_age = 14)
convert_animal_age(animal = "fish", human_age = 2)
```

# Functions with Data Bases

```{r}

#creatig a data frame with tribble, ~"name" = column names then hit enter and populate what will be in those coumns

dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

# within the function pulling information from the data frame to populate a statement, wrangling the data first

dog_menu <- function(name){
  my_sub <- filter(dog_choice, dog_name == name)
  print(paste0("My name is ", my_sub$dog_name, " and my favorite food is ", my_sub$food, "!"))
}

dog_menu("Banjo")
dog_menu("Waffle")

#still have a data frame but just 

# sort or filter whatever it was and pull in the information from the 

# first create a subset of the users using filter

```

Filtering Options
filter is the rows functions 

```{r}

# option 1

df |> 
  filter(condition)
#filter is built to know that if you feed some data into it, it assumes the first argument


# OPTION 2 but you can also specify the data with (data, condition)

filter(dataframe, condition)

```

# Functions Meet For Loops

```{r}

# write a function that iterates every variable in the functions prints out a mean and then pull out column names , iterating over columns of a data frame

means_over_columns <- function(df) {
  for (i in seq(df)){
    col_mean <- mean(df[[i]], na.rm= TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

# now using a data frame we can use the function to calculate the means

means_over_columns(df = mtcars)
means_over_columns(df = faithful)

# essentially creating the apply function
# calling columns from data frame need double brackets
# colnames is a specific function to print out column names
# col_mean calculates a the column means

```

