---
title: "Day 4 Notes and Comp Time"
author: "Colleen McCamy"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)

```

# Brain warm-up - creating a for loop

```{r}

# total time for any two quater miles; find all the times for any consecutive two quarters

quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

for (i in seq_along(quarter_splits)){
  print(quarter_splits[i] + quarter_splits [i+1])
}


# need to specific where the i is indexing from - thus include the vector[i]

```

# Intro to functions
making a function to double the user input

```{r}

double_this <- function(my_value){
  print(2 * my_value)
}

# now using the function lik

double_this(my_value = 14)

double_this(200)

# body of function tell it what to do

# don't need to specify argument name but need to have the same order

double_this(c(1, 2, 10, 100, 175))

#can use a vector too
```


Write a function that asks for a users name and age and returns "I'm [name] and I am [age] years old

```{r}

age_name <- function(name, age) {
  print(paste("I'm", name, "and I am", age, "years old." ))
}

age_name(name = "Colleen", age = 26)

# need to use paste where there are different types of data

```

# Looking at function documentation
Docs contains: 
- Title of function
- what the function name is and package in the top left
- description = brief summary
- usage = an idea of what the actual structure is of the function and what arguments are required for the function
- arguments = idea of default settings that are included
- Value = ideas of what outputs are from model functions; can show what all the different outputs are stored in that list

# Function with a conditional

```{r}

convert_animal_age <- function(animal, human_age){
  if (animal == "dog"){
    print(human_age *7)
  } else if (animal == "goat"){
    print(human_age * 4.7)
  } else {
    print("Please input a dog or goat")
    }
}

convert_animal_age(animal = "dog", human_age = 12)
convert_animal_age(animal = "goat", human_age = 14)
convert_animal_age(animal = "fish", human_age = 2)
```

# Functions with Data Bases

```{r}

#creatig a data frame with tribble, ~"name" = column names then hit enter and populate what will be in those coumns

dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

# within the function pulling information from the data frame to populate a statement, wrangling the data first

dog_menu <- function(name){
  my_sub <- filter(dog_choice, dog_name == name)
  print(paste0("My name is ", my_sub$dog_name, " and my favorite food is ", my_sub$food, "!"))
}

dog_menu("Banjo")
dog_menu("Waffle")

#still have a data frame but just 

# sort or filter whatever it was and pull in the information from the 

# first create a subset of the users using filter

```

Filtering Options
filter is the rows functions 

```{r}

# option 1

df |> 
  filter(condition)
#filter is built to know that if you feed some data into it, it assumes the first argument


# OPTION 2 but you can also specify the data with (data, condition)

filter(dataframe, condition)

```

# Functions Meet For Loops

```{r}

# write a function that iterates every variable in the functions prints out a mean and then pull out column names , iterating over columns of a data frame

means_over_columns <- function(df) {
  for (i in seq(df)){
    col_mean <- mean(df[[i]], na.rm= TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

# now using a data frame we can use the function to calculate the means

means_over_columns(df = mtcars)
means_over_columns(df = faithful)

# essentially creating the apply function
# calling columns from data frame need double brackets
# colnames is a specific function to print out column names
# col_mean calculates a the column means

```


# Example using a nested for loop with a matrix

```{r}

fish_color <- c("red", "blue")
fish_count <- c(1, 2)

fish_seus <- matrix(nrow = length(fish_color), ncol = length(fish_count))

for(i in seq_along(fish_color)){
  for (j in seq_along(fish_count)){
    color_count <- paste(fish_color[i], fish_count[j])
    fish_seus [j, i] <- color_count 
  }
}

fish_seus


```

Example - nested for loops part 2

```{r}

clothing <- c("shirt", "pants", "shoes")
color <- c("purple", "red", "blue", "white")

our_clothes <- matrix(nrow = length(color), ncol = length(clothing))

for(i in seq(clothing)){
  for (j in seq_along(color)){
    clothing_details <- paste(color[j], clothing[i])
    our_clothes[j,i] <- clothing_details
  }
}


our_clothes

```


# Shortcut for Functions

```{r}

# if you just write the right hand side of the function then do command + option + x 

my_shortcut <- function(oranges, rainbows) {
  (oranges * rainbows / 2.4)^2
}

```


# Logistic growth
Modeling a population over a time and varying growth rate

```{r}

logistic_growth <- function(K, N0, r, time) {
  K / (1 + ((K - N0) / N0) * exp(-r * time))
}

# Testing the function:

logistic_growth(K = 5000, N0 = 100, r = 0.07, time = 42)

```


```{r}

# Iterating over time now

time_vec <- seq(from = 0, to = 35, by = 0.1)

pop_35 <- logistic_growth(N0 = 40, time = time_vec, K = 8000, r = 0.3)

# created a seq of time and we have two separate vectors for time_vec and pop_35 - need to combine

pop_time35_df <- data.frame (time_vec, pop_35)

# data.frame will combine two vectors WITH the same length

```

Plotting the data frame of the change in population over time of the parameter of time

```{r}

ggplot(data = pop_time35_df, aes(x = time_vec, y = pop_35)) + geom_line(color = "pink") + labs(title = "Population Over Time")

# need to tell it three things, 1. that it is a ggplot, data and x and y axis, and what type of plot

```

# Another way to do this same thing with a for loop

```{r}

pop_35_vec <- vector(mode = "numeric", length = length(time_vec))

for (i in seq_along(time_vec)){
  population <- logistic_growth(N0 = 100, K = 6000, r =0.27, time = time_vec[i])
  pop_35_vec[i] <- population
}

```

# What if we wanted to check out the curves for multiple growth rates?
Now extending this over to iterate over a range of growth rates

```{r}
# creating a sequence for population growth rate

r_seq <- seq(from = 0.2, to = 0.4, by = 0.01)

```

```{r}
# first create empty matrix

out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

for (i in seq_along(r_seq)){
  for (j in seq_along(time_vec)){
    pop <- logistic_growth(N0 = 700, K = 80, r = r_seq[i], time = time_vec[j])
    out_matrix[j, i] <- pop
  }
}

# this populates the out matrix

```

# Now we have some wrangling to do

```{r}

# want to convert to data frame and bind it to the time vector to have one column with time and the subsequent population datas

logistic_df <- data.frame(out_matrix, time = time_vec)
# this adds a column time

colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")
# updating column names, this is looking in the r sequence and then create names with like gr_0.2; NEED same number of column names


# we now have a single variable for growth rate that is split up over multiple columns
# tidy data structure is needed - going to make it so that each variable only has a single column (we have three different variables - population in its own column, single column for population

# EVERY VARIABLE TO HAVE ITS OWN COLUMN

# Using Pivot Longer

logistic_df_long <- logistic_df |> 
  pivot_longer(cols = -time, values_to = "population", names_to = "growth_rate")

# we want to pivot longer for all of the populations except the time, first is what columns do you want to pivot longer? use excluding code to exclude time -time

# we want each column to show growth rate and then the actual value of the growth sequence that we made


```

#Plotting this now

```{r}

ggplot(data = logistic_df_long, aes(x = time, y = population)) + geom_line()
# this doesn't show the individual rate - seeing a very high frequenceline, want to tell the variable of the ascethic

ggplot(data = logistic_df_long, aes(x = time, y = population)) + geom_line(aes(color = growth_rate)) 

# this shows all of the different lines

```


# Making fucntion with warning and an error message

Use `warning` to add a warning.
Use `stop` to add an error.

Calcuate the power of wind hitting a turbine

$$ 
P=0.3\rho A v^3
$$
rho = air density (kg / m^3)
A = area covered by turbine (square meters)
v = wind velocity (m/s)

```{r}
# writing a function to calculate the collectable wind power with these three units

calc_windpower <- function(rho, radius, windspeed) {
  
  if (windspeed > 130) {
    warning("Whoa, that is really fast are you sure?")
  }
  
  if (rho > 1.225) {
    warning("That air density is uspicious.")
  }
  
  if (radius < 0) {
    stop("Rotor radius must be a positive value (meters).")
  }
  
  0.3*rho*pi*(radius^2)*(windspeed^3)
}


# checking this out
calc_windpower(rho = 1.225, radius = 50, windspeed = 35)
calc_windpower(rho = 1.225, radius = -6, windspeed = 15)
calc_windpower(rho = 3.2, radius = 10, windspeed = 15)

```



Never mistake a non-error as correct!


